Universals of word order result from optimization of grammars for efficient communication

### 2019-04-06 second-draft/3-paper-names.tex
revising

% ,lin2017critical
\dropcap{U}nderstanding and explaining the variation between human languages is the central  goal of Linguistics \cite{greenberg1963universals,chomsky1993lectures,givon2009genesis, ...}.
% behaghel1909beziehungen,
% \cite{behaghel1909beziehungen,zipf1949human,greenberg1963universals,lin2017critical} \cite{hawkins2007processing,bender2009linguistically,bender2013linguistic}
An explanation for the universal properties of language would enable a deeper scientific understanding of what human language is and how to model it, with applications in psychology and natural language processing \cite{hawkins2007processing,bender2009linguistically,bender2013linguistic}.
Across theoretical paradigms, linguists have hypothesized that language is shaped by efficiency in computation~\cite{chomsky2005three,hauser2002faculty, berwick1984grammatical,hawkins1994performance} and communication~\cite{zipf1949human, Croft:Cruse:2004, Goldberg:2005, pinker1990natural}.
%Role of general computational constraints ~\cite{hauser2002faculty} % berwick2016only
In this work we examine this question from a computational perspective, demonstrating a fully formalized and implemented framework in which a large class of language universals can be explained through the computational optimization of grammars for efficiency. % information-theoretic

@@

%We aim to explain the linguistic universals related to word order first documented by Greenberg \cite{greenberg1963universals}. 
Natural languages vary in the order in which they express information.
Consider Figure~\ref{fig:arabic-japanese-simple}, showing a sentence in Arabic (top) and Japanese (bottom), both translating to `I wrote a letter to a friend.'
Both sentences contain a verb meaning `wrote', a noun expressing  the object `letter', and an adpositional phrase translating to `to a friend'.
%In linguistic terminology, `letter' is the object of the verb, whereas the phrases translating as `to a friend' are known as adpositional phrases.
Yet, the order of these words are entirely different in the two languages:
%However, the languages differ in the order in which these words and phrases are arranged: 
The verb stands at the beginning in Arabic, and at the end in Japanese.
Arabic expresses `to' by a \emph{preposition} (\emph{preceding} the noun `friend'); Japanese uses a \emph{postposition} (\emph{following} it). % and a postposition in Japanese, following `friend'.
%In Arabic, the verb stands at the beginning, followed by the objecat and the adpositional phrase.
%In Arabic, `to' is expressed by a \emph{preposition}, so named because it precedes the word expressing `friend'. %The adpositional phrase starts with the adposition.
%In Japanese, it is expressed by a \textit{postposition}, so named because it follows the word denoting `friend'.
%What both languages have in common is that the object and the adpositional phrase stand on the same side of the verb, and that the objects of adpositions and verbs stand on the sam side of their heads.
%What both languages have in common is that the object and the adpositional phrase stand on the same side of the verb.
%It turns out that these are deep and stable regularities:

@@

This generalization falls in a group of language universals that were originally documented by Greenberg \cite{greenberg1963universals}, known as \key{word order correlations}:
These describe correlations between the relative positions of different types of expressions across languages.
The example above documents that the position of the object (`letter') relative to the verb is \key{correlated} with the position of the adposition (`to'). %adpositional phrases (`to a friend') relative to the verb.
Greenberg also found that the order between verb and object is correlated with other aspects of a language's word order, such the order of verb and adpositional phrase (`wrote -- to friend' in Arabic vs `friend to -- wrote' in Japanese), and that of noun and genitive (`book -- of friend' in Arabic, `friend of -- book' in Japanese).
Greenberg originally worked with 30 languages; the correlation universals have since been confirmed on the basis of much larger samples of languages.
The authoritative study by Dryer \cite{dryer1992greenbergian} draws on 625 languages and documents fifteen such correlations.

@@

@@

Recent work has used an information-theoretic formalization to computationally test these ideas in various domains of language, showing that they predict
both basic statistical properties of languages~\cite{ferreri2003least},
and sophisticated aspects of language, such as pragmatic inference \cite{frank2012predicting}, and the distribution of color words~\cite{zaslavsky2018efficient} and kinship categories~\cite{kemp2012kinship} across many languages.
Common to these accounts is the idea of maximizing the amount of information that linguistic forms provide about meanings, while constraining the diversity of forms.

@@

@@

\begin{figure}
    \centering
    \includegraphics[width=0.45\textwidth]{figures/communication-two-langs.png}
	\caption{Grammars define communication channels: A speaker (left) expresses a dependency structure into a string of words forming a sentence. The order of words is determined by the grammar of the language. A listener recovers a dependency structure.}
	\label{fig:comm}
\end{figure}

@@

Consider Figure~\ref{fig:comm}.
Imagine the speaker wants to communicate that Mary saw John.
The message (left) consists of a dependency structure with the words `saw', `Mary', `John', where Mary is the subject of the event denoted by `saw', and John is the object.
When uttering, the speaker needs to choose an ordering in which to order the words in this tree to generate a string of words.
In natural language, this order is largely determined by the \key{grammar} of the language (CITE).
The speaker will order it as \emph{Mary saw John} (Figure~\ref{fig:comm}, top) in English, and as \emph{Mary John saw} (Figure~\ref{fig:comm}, bottom) in Japanese.
%
%(1)
%\begin{dependency}[theme=simple]
%    \begin{deptext}
%    Mary \& saw \& John \\
%    \end{deptext}
%    \depedge{2}{1}{nsubj}
%    \depedge{2}{3}{obj}
%\end{dependency}
%(2)
%\begin{dependency}[theme=simple]
%    \begin{deptext}
%    Mary \& John  \& saw \\
%    \end{deptext}
%    \depedge{3}{1}{nsubj}
%    \depedge{3}{2}{obj}
%\end{dependency}
%
The listener receives a sentence -- a linear string of words -- and decodes the underlying message in accordance with the grammar of the language (Figure~\ref{fig:comm}, right).
%Consistent with Shannon's model, we assume a Bayesian listener who decodes dependency structures probabilistically~(Figure~\ref{fig:comm}).
%If communication is successful, the listener can identify the intended structure with high confidence~(Figure~\ref{fig:comm}).

@@

%All natural languages have some degree of word order regularities that are specified in their grammar.
%For instance, English places subjects before and objects after the verb, allowing the listener to unambiguously decode the sentence in Figure~\ref{fig:comm}.
%That is, speakers and listeners have shared knowledge of a \key{grammar} that specifies how dependency structures are encoded into sentences.
%Natural languages differ in the rules they apply: Some place the object after the verb, some place it before the verb.
%Speakers encode dependency trees into sentences according to a grammar~(Figure~\ref{fig:grammars}).
%While natural languages differ in the freedom of word order, all languages rely on some amount of word order reglarity to encode meaning.
The grammars of natural languages provide consistent ordering rules for all syntactic relations.
This is illustrated in Figure~\ref{fig:grammars}:
We show how four different grammars order objects, adpositional phrases, and adpositions.
For instance, Grammar 1 -- corresponding to Arabic in Figure~\ref{fig:arabic-japanese-simple} -- orders objects (`friends', `letter') after verbs and has prepositions (`to friend').
Grammar 2 orders objects after verbs but has postpositions (`friend -- to').
Grammars 3 and 4 place the object before the verb, and one of them (Grammar 3) corresponds to Japanese order.

@@

%In information-theoretic formalizations of Zipf's ideas, the efficiency of a 
In Information Theory, the usefulness of a communication channel depends on (1) how costly encoding and transmission are, and (2) how precisely messages can be recovered from codes~\cite{shannon1948mathematical}.
%This idea formalizes Zipf's forces of Unification and Diversification
Overall efficiency can be formalized as a weighted combination of terms representing the amount of information conveyed and the cost of communication:
%The \key{efficiency} of a language is a weighted combination
\begin{equation}\label{eq:efficiency}
	R_{\textit{Eff}} := R_{\textit{Informativity}} - \lambda R_\textit{Cost}
\end{equation}
with an interpolation weight $\lambda \in [0,1)$. %In all experiments in this paper we use $\lambda=.9$ (see SI appendix section 5 for mathematical justification and for robustness to other choices).
This objective function has been proposed in prior computational tests of functional explanations~\cite{ferreri2003least,frank2012predicting,zaslavsky2018efficient,kemp2012kinship}.

@@

In our model, both aspects of efficiency will depend on the grammar:
Grammars can differ in the degree to which a listener can unambiguously recover the dependency structure from an utterance.
This reflects the \key{parseability} of the language.
Maximizing the information provided by utterances about meanings is a standard formalization of Zipf's Force of Diversification~\cite{frank2012predicting,regier2015word}. % also goodman2013knowledge,
The degree to which listeners can reconstruct dependency structures from an utterance is formalized as the amount of information that utterances $u$ provide about their underlying dependency structures $t$:
\begin{equation}
	R_{Pars} := I[\mathcal{U},\mathcal{T}] = \sum_{t,u} p(t,u) \log \frac{p(t|u)}{p(t)}
\end{equation}
where the sum runs over all possible pairs of sentences $l$ and dependency structures $t$ in the language.
This quantity describes the degree to which dependency structures can be unambiguously recovered from sentences.
It is large when they can mostly be recovered, and small if sentences have high amounts of ambiguity about the dependency structure.
In line with Zipf's Force of Diversification, this quantity is maximized when trees can be decoded fully unambiguously from utterances.

We formalize the cost of an utterance as its word-by-word surprisal, the degree to which it is unpredictable from the general statistics of the language.
Surprisal has been found to be a highly accurate and general predictor human online processing difficulty \cite{hale2001probabilistic,levy2008expectation,smith2013effect}, and can be justified  in terms of predictive coding theories of information processing in the brain \cite{friston2009predictive} as well as the general algorithmic complexity of language generation and comprehension \cite{li2008introduction}.
In expectation over all utterances $u$ in a language, the negative surprisal describes the \key{predictability}, or negative entropy, of the utteranes: %this quantity is the entropy: % of the sentences:
%The \key{predictability} of the language is the negative entropy of the sentences:
\begin{equation}
	R_{Pred} := - H[\mathcal{U}] = \sum_{u} p(u) \log p(u)
\end{equation}
where the sum runs over all possible sentences $l$ that belong to the language.
In keeping with Zipf's Force of Unification, this quantity describes how homogeneous the language is, i.e., it is larger if the distribution over sentences is concentrated on a smaller number of frequent sentences. %how many distinct sentences there are.

With these choices, our objective function is equivalent to that proposed in \cite{ferrericancho2002zipf,ferrericancho2007global,futrell2017memory} as a general objective for communicative systems. %, taking dependency structures as the underlying meanings to be conveyed.
These works mainly differ in the cost function applied: can use the number of different utterances (CITE), the complexity of meanings \cite{kemp2012kinship}, and the entropy of possible utterances \cite{ferrericancho2002zipf,ferrericancho2007global,futrell2017memory}.

@@

We examined 
(1) whether the grammars of natural languages evolve towards optimizing efficiency of communication, and
(2) whether this process of optimization accounts  for Greenberg's order correlation universals.
%We formalize word order grammars as specifying how different types of dependents (subjects, objects, etc.) are ordered, employing a cross-linguistically meaningful inventory of relations~\cite{nivre2015universal}.

@@

%
%

@@

		& \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-ground-lifted_case.pdf} } %\textbf{86}  }
	%	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-langmod-lifted_case.pdf} } %  \textbf{47 \%}$^{}$   }
		&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-efficiency_large_best-lifted_case.pdf}     } & \\
		&  \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-ground-lifted_cop.pdf} }% \textbf{94} }
		&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-efficiency_large_best-lifted_cop.pdf}   } & \\
		&  \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-ground-aux.pdf}}
		&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-efficiency_large_best-aux.pdf}  } & \\
		&  \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-ground-nmod.pdf}}
		&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-efficiency_large_best-nmod.pdf}  } & \\
		&  \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-ground-acl.pdf}}
		&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-efficiency_large_best-acl.pdf}  } & \\
		&  \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-ground-lifted_mark.pdf}}
		&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-efficiency_large_best-lifted_mark.pdf}  }  &\\
		&  \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-ground-obl.pdf}}
		&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-efficiency_large_best-obl.pdf}  } &  \\
		&  \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-ground-xcomp.pdf}}
		&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-efficiency_large_best-xcomp.pdf}  } & \\
\caption{OPTION 1: Efficiency optimization accurately predicts the Greenbergian correlations. Each correlation is stated in terms of a pair of a `verb patterner' and an `object patterner', whose relative order correlate with that of verbs and objects.

@@

	%	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-langmod-lifted_case.pdf} } %  \textbf{47 \%}$^{}$   }

@@

@@

@@

\begin{table*}
\begin{tabular}{c|ll|c|c|cc|ccccccccccc}
	&	\multicolumn{2}{|c|}{Correlates with...}   &         \multirow{2}{*}{Real}   &  \multirow{2}{*}{Efficiency} &  \multirow{2}{*}{Predictability} &  \multirow{2}{*}{Parseability} & \\ 
		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_lifted_case.pdf}     }
		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Predictability_lifted_case.pdf}     } 
		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Parseability_lifted_case.pdf}     }  & \\
		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_lifted_cop.pdf}     } 
		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Predictability_lifted_case.pdf}     }
	&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Parseability_lifted_cop.pdf}     } 
& \\
	&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_aux.pdf}     } 
	&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Predictability_aux.pdf}     } 
		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Parseability_aux.pdf}     } & \\
	&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_nmod.pdf}     } 
	&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Predictability_nmod.pdf}     } 
		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Parseability_nmod.pdf}     } & \\
	&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_lifted_cop.pdf}     }
	&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Predictability_lifted_cop.pdf}     }
		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Parseability_acl.pdf}     } & \\
		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_lifted_mark.pdf}     } 
		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Predictability_lifted_mark.pdf}     } 
		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Parseability_lifted_mark.pdf}     } & \\
		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_lifted_mark.pdf}     } 
		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Predictability_lifted_mark.pdf}     } 
		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Parseability_obl.pdf}     } & \\
		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_xcomp.pdf}     }
		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Predictability_xcomp.pdf}     }
		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Parseability_xcomp.pdf}     } & \\
\caption{OPTION 3: Efficiency optimization accurately predicts the Greenbergian correlations. Each correlation is stated in terms of a pair of a `verb patterner' and an `object patterner', whose relative order correlate with that of verbs and objects.
\end{table*}

@@

The eight correlations are shown in Table~\ref{table:corr-dryer}.

We compare the prevalence of the word order correlations in optimized languages (right) to their prevalence in the real languages (left).

For the real languages, we indicate how many of the 51 languages satisfy a correlation.
For the optimized languages, we indicate the posterior distribution of the proportion of satisfying languages, obtained from a mixed-effects logistic regression with language families as random effects (see \textit{Materials and Methods}).
Table~\ref{table:corr-dryer} shows that grammars optimized for efficiency accurately predict all eight correlations to hold at prevalences similar to actual languages.

>>>>

% ,lin2017critical
\dropcap{U}nderstanding and explaining the variation between human languages is the central  goal of Linguistics \cite{greenberg1963universals,chomsky1993lectures,givon2009genesis, ...}.
% behaghel1909beziehungen,
% \cite{behaghel1909beziehungen,zipf1949human,greenberg1963universals,lin2017critical} \cite{hawkins2007processing,bender2009linguistically,bender2013linguistic}
An explanation for the universal properties of language would enable a deeper scientific understanding of what human language is and how to model it, with applications in psychology and natural language processing \cite{hawkins2007processing,bender2009linguistically,bender2013linguistic}.
Across theoretical paradigms, linguists have hypothesized that language is shaped by efficiency in computation~\cite{chomsky2005three,hauser2002faculty, berwick1984grammatical,hawkins1994performance} and communication~\cite{zipf1949human, Croft:Cruse:2004, Goldberg:2005, pinker1990natural, smith2013linguistic}.
%Role of general computational constraints ~\cite{hauser2002faculty} % berwick2016only
In this work we examine this question from a computational perspective, demonstrating a fully formalized and implemented framework in which a large class of language universals can be explained through the computational optimization of grammars for efficiency. % information-theoretic

@@

%We aim to explain the linguistic universals related to word order first documented by Greenberg \cite{greenberg1963universals}. 
Human languages vary in the order in which they express information.
Consider Figure~\ref{fig:arabic-japanese-simple}, showing a sentence in Arabic (top) and Japanese (bottom), both translating to `I wrote a letter to a friend.'
Both sentences contain a verb meaning `wrote', a noun expressing  the object `letter', and an adpositional phrase translating to `to a friend'.
%In linguistic terminology, `letter' is the object of the verb, whereas the phrases translating as `to a friend' are known as adpositional phrases.
Yet, the order of these words are entirely different in the two languages:
%However, the languages differ in the order in which these words and phrases are arranged: 
The verb stands at the beginning in Arabic, and at the end in Japanese.
Arabic expresses `to' by a \emph{preposition} (\emph{preceding} the noun `friend'); Japanese uses a \emph{postposition} (\emph{following} it). % and a postposition in Japanese, following `friend'.
%In Arabic, the verb stands at the beginning, followed by the objecat and the adpositional phrase.
%In Arabic, `to' is expressed by a \emph{preposition}, so named because it precedes the word expressing `friend'. %The adpositional phrase starts with the adposition.
%In Japanese, it is expressed by a \textit{postposition}, so named because it follows the word denoting `friend'.
%What both languages have in common is that the object and the adpositional phrase stand on the same side of the verb, and that the objects of adpositions and verbs stand on the sam side of their heads.
%What both languages have in common is that the object and the adpositional phrase stand on the same side of the verb.
%It turns out that these are deep and stable regularities:

@@

This generalization falls in a group of language universals that were originally documented by Greenberg \cite{greenberg1963universals}, known as \key{word order correlations}:
These describe correlations between the relative positions of different types of expressions across languages.
The example above documents that the position of the object (`letter') relative to the verb is \key{correlated} with the position of the adposition (`to'). %adpositional phrases (`to a friend') relative to the verb.
Greenberg also found that the order between verb and object is correlated with other aspects of a language's word order (Table~\ref{table:corr-dryer}), such the order of verb and adpositional phrase (`wrote -- to friend' in Arabic vs `friend to -- wrote' in Japanese), and that of noun and genitive (`book -- of friend' in Arabic, `friend of -- book' in Japanese).
%Greenberg originally worked with 30 languages; the correlation universals have since been confirmed on the basis of much larger samples of languages.
%The authoritative study by Dryer \cite{dryer1992greenbergian} draws on 625 languages and documents fifteen such correlations.

@@

% alternatively: expressivity vs simplicity

@@

Recent work has used an information-theoretic formalization to computationally test these ideas in various domains of language, showing that they predict
both basic statistical properties of languages~\cite{ferreri2003least} and language development~\cite{smith2013linguistic},
and sophisticated aspects of language, such as pragmatic inference \cite{frank2012predicting}, and the distribution of color words~\cite{zaslavsky2018efficient} and kinship categories~\cite{kemp2012kinship} across many languages.
Common to these accounts is the idea of maximizing the amount of information that linguistic forms provide about meanings, while constraining the diversity of forms.

@@

\begin{table}
	\begin{center}
%		\setlength\tabcolsep{1.5pt}
\begin{tabular}{|c|ll|ll|}
	\hline
	%&
%	&	\multicolumn{2}{|c|}{Correlates with...}   &         \multirow{2}{*}{Real}   &  \multirow{2}{*}{Predicted} & \\ 
%	&	Verb & Object     & Object & Verb    \\ \hline\hline
	& kataba & rasaa'il 	&tegami-o & kaita \\
	&	\emph{wrote} & \emph{letters} & 	\emph{letter} & \emph{wrote} \\ \hline \hline
	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {1}}}}	&	li    &    sadiiq       &	tomodachi & ni \\
	&		\emph{to}            & \emph{a friend} &		\emph{friend} & \emph{to} \\ \hline
	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {2}}}}	&kaana    &    sadiiq         &	tomodachi & datta \\
	&	\emph{was}        & \emph{a friend} 	&\emph{friend} & \emph{was} \\ \hline
	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {3}}}}	&sawfa    &    yaktub       & 	kak & udesho \\
	&	\emph{will}          & \emph{write}  &    	\emph{write} & \emph{will} \\ \hline
	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {4}}}}	& sadiiq  &    John    & 	John no & tomodachi \\
	&	\emph{friend} &  \emph{of John}  &	\emph{John of} & \emph{friend} \\ \hline
	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {5}}}}	&kutub    &    taqra'uha       & 	anata-ga yonda & hon \\
	&	\emph{books} & \emph{that you read}  &	\emph{that you read} & \emph{book} \\ \hline
	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {6}}}}	&'an    &    tusil        & 	toochaku suru & koto \\
	&	\emph{that} & \emph{she arrives}  &	\emph{has.arrived} & \emph{that} \\ \hline
	\multirow{2}{*}{	\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {7}}}}	&dhahabt	    &    'ila lmadrasa     & 	gakkoo ni & ittekimashita \\
	&	\emph{went} & \emph{to school}  &	\emph{school to} & \emph{went} \\ \hline
	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {8}}}}	&'uriid    &    'an 'ughaadir         & 	yom & itai \\
	& \emph{wants}   &  \emph{to leave}  &	\emph{to leave} & \emph{want} \\ \hline \hline
%comp. &
%\multirow{3}{*}{NP}&
%	AP &
%    \multicolumn{4}{l}{\footnotesize{Significance levels: $^*$: $p < 0.05$, : $p < 0.01$, : $p < 0.001$}}
\end{tabular}
	\end{center}
	\caption{Greenberg's word order correlations, exemplified by Arabic (left) and Japanese (right) examples. Based on Dryer \cite{dryer1992greenbergian}, see Materials and Methods.}\label{table:corr-dryer}
\end{table}

@@

\begin{figure}
    \centering
    \includegraphics[width=0.45\textwidth]{figures/communication-two-langs.png}
	\caption{Grammars define communication channels: A speaker (left) expresses a dependency structure into a string of words forming a sentence. The order of words is determined by the grammar of the language. A listener recovers a dependency structure. Communication is successful if the listener recovers the correct depenency structure.}
	\label{fig:comm}
\end{figure}

@@

Consider Figure~\ref{fig:comm}.
Imagine the speaker wants to communicate that Mary saw John.
The message (left) consists of a dependency structure with the words `saw', `Mary', `John', where Mary is the subject of the event denoted by `saw', and John is the object.
When uttering, the speaker needs to choose an ordering in which to order the words in this tree to generate a string of words.
In human language, this order is largely determined by the \key{grammar} of the language (CITE).
The speaker will order it as \emph{Mary saw John} (Figure~\ref{fig:comm}, top) in English, and as \emph{Mary John saw} (Figure~\ref{fig:comm}, bottom) in Japanese.
%
%(1)
%\begin{dependency}[theme=simple]
%    \begin{deptext}
%    Mary \& saw \& John \\
%    \end{deptext}
%    \depedge{2}{1}{nsubj}
%    \depedge{2}{3}{obj}
%\end{dependency}
%(2)
%\begin{dependency}[theme=simple]
%    \begin{deptext}
%    Mary \& John  \& saw \\
%    \end{deptext}
%    \depedge{3}{1}{nsubj}
%    \depedge{3}{2}{obj}
%\end{dependency}
%
The listener receives a sentence -- a linear string of words -- and decodes the underlying message in accordance with the grammar of the language (Figure~\ref{fig:comm}, right).
%Consistent with Shannon's model, we assume a Bayesian listener who decodes dependency structures probabilistically~(Figure~\ref{fig:comm}).
%If communication is successful, the listener can identify the intended structure with high confidence~(Figure~\ref{fig:comm}).

@@

%All natural languages have some degree of word order regularities that are specified in their grammar.
%For instance, English places subjects before and objects after the verb, allowing the listener to unambiguously decode the sentence in Figure~\ref{fig:comm}.
%That is, speakers and listeners have shared knowledge of a \key{grammar} that specifies how dependency structures are encoded into sentences.
%Natural languages differ in the rules they apply: Some place the object after the verb, some place it before the verb.
%Speakers encode dependency trees into sentences according to a grammar~(Figure~\ref{fig:grammars}).
%While natural languages differ in the freedom of word order, all languages rely on some amount of word order reglarity to encode meaning.
The grammars of human languages provide consistent ordering rules for all syntactic relations.
This is illustrated in Figure~\ref{fig:grammars}:
We show how four different grammars order objects, adpositional phrases, and adpositions.
For instance, Grammar 1 -- corresponding to Arabic in Figure~\ref{fig:arabic-japanese-simple} -- orders objects (`friends', `letter') after verbs and has prepositions (`to friend').
Grammar 2 orders objects after verbs but has postpositions (`friend -- to').
Grammars 3 and 4 place the object before the verb, and one of them (Grammar 3) corresponds to Japanese order.

@@

%Prior computational tests of efficiency explanations for various aspects of language have

%In information-theoretic formalizations of Zipf's ideas, the efficiency of a

%In Information Theory, the usefulness of a communication channel 
The efficiency of any information-theoretic communication protocol depends on (1) how costly encoding and transmission are, and (2) how precisely messages can be recovered from codes~\cite{shannon1948mathematical}.
%This idea formalizes Zipf's forces of Unification and Diversification
Accordingly, computational tests of functional explanations~\cite{ferreri2003least,frank2012predicting,zaslavsky2018efficient,kemp2012kinship,regier2015word,goodman2013knowledge} have formalized the overall efficiency of language as a weighted combination of terms representing the amount of information that utterances contain about the underlying messages, and the cost of communication:
%The \key{efficiency} of a language is a weighted combination
\begin{equation}\label{eq:efficiency}
	R_{\textit{Eff}} := R_{\textit{Informativity}} - \lambda R_\textit{Cost}
\end{equation}
with an interpolation weight $\lambda \in [0,1)$. %In all experiments in this paper we use $\lambda=.9$ (see SI appendix section 5 for mathematical justification and for robustness to other choices).

@@

%In our model, both aspects of efficiency will depend on the grammar:
Grammars differ in the degree to which a listener can unambiguously recover the dependency structure from an utterance.
This reflects the \key{parseability} of the language.
%Maximizing the information provided by utterances about meanings is a standard formalization of Zipf's Force of Diversification~\cite{frank2012predicting,regier2015word}. % also goodman2013knowledge,
The degree to which listeners can reconstruct dependency structures from an utterance is formalized as the amount of information that utterances $u$ provide about their underlying dependency structures $t$:
\begin{equation}
	R_{Pars} := I[\mathcal{U},\mathcal{T}] = \sum_{t,u} p(t,u) \log \frac{p(t|u)}{p(t)}
\end{equation}
where the sum runs over all possible pairs of sentences $l$ and dependency structures $t$ in the language.
This quantity describes the degree to which dependency structures can be unambiguously recovered from sentences.
%It is large when they can mostly be recovered, and small if sentences have high amounts of ambiguity about the dependency structure.
%In line with Zipf's Force of Diversification, it is maximized when trees can be decoded fully unambiguously from utterances.

%This idea formalizes Zipf's forces of Unification and Diversification
%Accordingly, computational tests of functional explanations~\cite{ferreri2003least,frank2012predicting,zaslavsky2018efficient,kemp2012kinship,regier2015word,goodman2013knowledge} have formalized the overall efficiency of language as a weighted combination of terms representing the amount of information that utterances contain about the underlying messages, and the cost of communication:

We formalize the cost of a language as its entropy~\cite{ferreri2003least,ferrericancho2007global,futrell2017memory}.
This is the average of the word-by-word surprisal, the degree to which it is unpredictable from the general statistics of the language.
Surprisal has been found to be a highly accurate and general predictor human online processing difficulty \cite{hale2001probabilistic,levy2008expectation,smith2013effect}, and can be justified  in terms of predictive coding theories of information processing in the brain \cite{friston2009predictive} as well as the general algorithmic complexity of language generation and comprehension \cite{li2008introduction}.
In expectation over all utterances $u$ in a language, the negative surprisal describes the \key{predictability}, or negative entropy, of the utteranes: %this quantity is the entropy: % of the sentences:
%The \key{predictability} of the language is the negative entropy of the sentences:
\begin{equation}
	R_{Pred} := - H[\mathcal{U}] = \sum_{u} p(u) \log p(u)
\end{equation}
where the sum runs over all possible sentences $l$ that belong to the language.
In keeping with Zipf's Force of Unification, this quantity describes how homogeneous the language is, i.e., it is larger if the distribution over sentences is concentrated on a smaller number of frequent sentences. %how many distinct sentences there are.

%With these choices, our objective function is equivalent to that proposed in \cite{ferrericancho2002zipf,ferrericancho2007global,futrell2017memory} as a general objective for communicative systems. %, taking dependency structures as the underlying meanings to be conveyed.
%These works mainly differ in the cost function applied: can use the number of different utterances (CITE), the complexity of meanings \cite{kemp2012kinship}, and the entropy of possible utterances \cite{ferrericancho2002zipf,ferrericancho2007global,futrell2017memory}.

@@

We examined 
(1) whether the grammars of human languages evolve towards optimizing efficiency of communication, and
(2) whether this process of optimization accounts  for Greenberg's order correlation universals.
%We formalize word order grammars as specifying how different types of dependents (subjects, objects, etc.) are ordered, employing a cross-linguistically meaningful inventory of relations~\cite{nivre2015universal}.

@@

@@

\begin{tabular}{|c|ll|c|cc|ccc}

@@

@@

\caption{Efficiency optimization accurately predicts the Greenbergian correlations.
	%Each correlation is stated in terms of a pair of a `verb patterner' and an `object patterner', whose relative order correlate with that of verbs and objects.
For each correlation, we provide the distribution among the actual grammars of the 51 languages (left), and among grammars optimized for efficiency (right).
%Blue bars count grammars conforming to the correlation, red bars count grammars violating it.

@@

}\label{table:corr-resu}

@@

@@

We first evaluated which of the eight correlations emerge from efficiency optimization across language families.
We controlled for variation across different optima by creating eight optimized grammars for each language.
We controlled for variation between languages and language families by constructing a single Bayesian mixed-effects logistic regression model predicting which correlations an optimized grammar satisfies, entering the language and language family as random effects.
In this analysis, efficiency optimization predicts all eight correlations to hold across languages with posterior probability $0.9911$.

In Table~\ref{table:corr-resu}, we compare the prevalence of the eight correlations in real and optimized languages.
For the real languages, we indicate how many of the 51 languages satisfy a correlation.
For the optimized languages, we indicate the posterior distribution of the proportion of satisfying languages, obtained from the mixed-effects analysis (see \textit{Materials and Methods}).
Grammars optimized for efficiency accurately predict all eight correlations to hold at prevalences significantly greater than 50\%, similar to actual human languages.
Optimizing for only predicability or only parseability does not predict all of the correlations (see SI).
